---
title: "Get started with liteformats"
output:
  html:
    meta:
      css: default
vignette: >
  %\VignetteIndexEntry{Get started with liteformats}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

liteformats provides lightweight document formats for
[litedown](https://github.com/yihui/litedown).
Source documents use Markdown with YAML metadata and can contain R code.
The rendered output is self-contained HTML or a PDF printed from that HTML.

# Choose a format

Each format has a starter, a renderer, and an options helper. Keep these
functions together when creating and rendering a document.

## Resume

Copy the resume starter, edit its YAML and Markdown, and render either output:

```{r}
#| eval = FALSE
liteformats::use_resume("resume.Rmd")
liteformats::resume("resume.Rmd", "resume.html")
liteformats::resume("resume.Rmd", "resume.pdf")
```

See `vignette("resume")` for the format's metadata and entry syntax.

## Cover letter

Use the cover letter starter with its matching renderer:

```{r}
#| eval = FALSE
liteformats::use_cover_letter("cover-letter.Rmd")
liteformats::cover_letter("cover-letter.Rmd", "cover-letter.html")
liteformats::cover_letter("cover-letter.Rmd", "cover-letter.pdf")
```

See `vignette("cover-letter")` for its metadata and fixed two-column composition.

# Choose an output

The output extension selects the result. HTML output is self-contained and
does not need a browser. PDF output first creates the same HTML, then prints
it with `xfun::browser_print()`. A local Chromium-based browser is
therefore required for PDF output only.

If `output` is omitted, the renderer writes an HTML file beside the source.
Both renderers return the normalized output path invisibly.

# Continue with the guides

- `vignette("config")` explains YAML, R options, and precedence.
- `vignette("typography")` covers type, spacing, font sources, and reliable text extraction.
- `vignette("workflows")` shows a small `Make` workflow for repeatable HTML and PDF builds.
