---
title: "Typography"
output:
  html:
    meta:
      css: default
vignette: >
  %\VignetteIndexEntry{Typography}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
---

The main typography and vertical rhythm controls can be set together in YAML:

``` yaml
liteformats:
  font-family: "Charter, Georgia, serif"
  font-size: 11pt
  font-scale: 0.94
  ligatures: false
  line-height: 1.15
  paragraph-spacing: 0.05em
  section-spacing: 1em
```

`font-scale` adjusts the effective font size without changing the nominal
`font-size`, which is useful when switching between families with different
metrics. The spacing controls accept CSS values. Format defaults may differ to
fit their content and composition.

# Use local fonts

A local family can provide regular, italic, bold, and bold italic font files.
OTF, TTF, WOFF, and WOFF2 files are embedded as data URIs, keeping the HTML
portable and allowing PDF rendering to work offline.

```{r}
#| eval = FALSE
liteformats::resume(
  "resume.Rmd",
  options = liteformats::resume_options(
    font_files = c(
      regular = "fonts/MySerif-Regular.otf",
      italic = "fonts/MySerif-Italic.otf",
      bold = "fonts/MySerif-Bold.otf"
    ),
    font_scale = 0.9,
    ligatures = FALSE
  )
)
```

Relative paths are resolved from the source document. The same typography
arguments are available in `cover_letter_options()`.

# Opt in to Google Fonts

Set `google-font: "Source Serif 4"` in YAML or `google_font = "Source Serif 4"`
in an options helper to add a Google Fonts stylesheet.
liteformats uses its bundled catalog to request available weights and styles
for regular, bold, italic, and bold italic text. It prefers a variable weight
range when the family provides one and omits unsupported combinations.

This is deliberately opt-in because it requires internet access when the
document is rendered. A full Google Fonts CSS v2 URL overrides the
catalog-derived request and can select explicit weights or variable axes:

```{r}
#| eval = FALSE
liteformats::resume(
  "resume.Rmd",
  options = liteformats::resume_options(
    google_font = paste0(
      "https://fonts.googleapis.com/css2?",
      "family=Inter:ital,opsz,wght@0,14..32,100..900;",
      "1,14..32,100..900&display=swap"
    )
  )
)
```

The full URL form also supports families newer than the bundled catalog.

# Check fitting and text extraction

Ligatures are disabled by default for more predictable text extraction from
PDFs. An embedded font's own tables can also affect extraction, so check the
final PDF when machine-readable text matters. For example:

``` sh
pdftotext document.pdf -
```

After changing a font, tune `font-scale`, `line-height`, paragraph and section
spacing, or page margins, then inspect both HTML and PDF.
See `vignette("config")` for page settings and configuration precedence.
